---
title: "repdata-010-project1"
author: "Youssef Shahin"
date: "January 17, 2015"
output: html_document
---

Activity data loading using `read.csv`
```{r}
library(lattice)

activity <- read.csv("activity.csv", header = T)
activity$date <- as.Date(activity$date)
activity$interval <- as.integer(activity$interval)
```

What is mean total number of steps taken per day?
```{r}
totals <- aggregate(steps ~ date, activity, sum, na.rm=T)
histogram(~steps, totals)
summary(totals)
```

What is the average daily activity pattern?
```{r}
totals.interval <- aggregate(steps ~ interval, activity, sum, na.rm=T)
xyplot(steps ~ interval, data=totals.interval, main = "average steps per interval", ylab = "steps", xlab = "interval", type ="l")
max_interval <- totals.interval$interval[which(totals.interval$steps == max(totals.interval$steps))]
print(paste0("Interval with max number of steps accross all days is ", max_interval))
```

Histogram after imputing NAs by filling in average value for time interval

```{r}
print(paste0("Number of rows with mising NAs is ", sum(!complete.cases(activity))))
activity.full <- activity[,]
activity.full[!complete.cases(activity.full),'steps'] <- mean(activity$steps, na.rm=T)
totals <- aggregate(steps ~ date, activity.full, sum)
histogram(~steps, totals)
summary(totals)
```

Using imputed NAs with mean of steps shows minor changes to the mean compared the the original value, while the median stayed the same.

Are there differences in activity patterns between weekdays and weekends?

```{r}
activity.type <- cbind(activity.full, type=as.factor(
  ifelse(weekdays(activity.full$date) == "Saterday" 
         | weekdays(activity.full$date) == "Sunday", "weekend", "weekday")))

totals <- aggregate(steps ~ interval + type, activity.type, mean)

xyplot(steps ~ interval | type, type = "l", layout = c(1, 2),
    main = "Average Number of Steps by Time Interval", data=totals,
    xlab = "Time Interval (5-mins)", ylab = "Average number of steps")
```
