---
title: "repdata-010-project1"
author: "Youssef Shahin"
date: "January 17, 2015"
output: html_document
---

Activity data loading using `read.csv`
```{r}
activity <- read.csv("activity.csv", header = T)
activity$date <- as.Date(activity$date)
activity$interval <- as.integer(activity$interval)
```

What is mean total number of steps taken per day?
```{r}
totals.daily <- tapply(activity$steps, activity$date, sum, na.rm=T)
barplot(totals.daily, main = "total steps per day", ylab = "steps", xlab = "day")
print(paste0("Mean of total number of steps per day is ", mean(totals.daily)))
print(paste0("Median of total number of steps per day is ", median(totals.daily)))
```

What is the average daily activity pattern?
```{r}
totals.interval <- tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(row.names(totals.interval), totals.interval, main = "average steps per interval", ylab = "steps", xlab = "interval", type ="l")
print(paste0("Interval with max number of steps accross all days is  ", names(sort(totals.interval, decreasing = T)[1])))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
print(paste0("Number of rows with mising NAs is ", sum(!complete.cases(activity))))
activity.interval.full <- activity[,]
activity.interval.full[!complete.cases(activity),]$steps <- totals.interval[as.character(activity.interval.full[!complete.cases(activity),]$interval)]

activity.daily.full <- activity[,]
activity.daily.full[!complete.cases(activity),]$steps <- totals.daily[as.character(activity.daily.full[!complete.cases(activity),]$interval)]

totals.daily.full <- tapply(activity.daily.full$steps, activity.daily.full$date, sum)
totals.daily.full <- tapply(activity.interval.full$steps, activity.daily.full$date, sum)
barplot(totals.daily.full, main = "total steps per day (w/Imputed NAs)", ylab = "steps", xlab = "day")

plot(totals.daily, col='blue')
points(totals.daily.full, col='red')

print(paste0("Mean of total number of steps per day w/imputed NAs is ", mean(totals.daily.full)))
print(paste0("Median of total number of steps per day w/imputed NAs is ", median(totals.daily.full)))
```

Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.